Physcraper: Body

### 2020-07-02 2-body.Rmd
comments
kinsey comments

More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007). With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt). Yet, publicly available estimates of phylogenetic relationships still cover only about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago. Some of this discrepance is because that many phylogenies are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay]. However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.

@@

@@

- The software MUSCLE [@edgar2004muscle] is used by default to perform sequence alignments.
- First, all new sequences are aligned using default MUSCLE options.
- Then, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
- The final alignment is not further processed automatically. So, we encourage users
  to check it either by eye and perform manual refinement or using any of the many
  tools for alignment processing, to eliminate columns with no information.
- Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.

>>>>

More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007). With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turn covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt). Yet, we only have data on phylogenetic relationships available for about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago. Some of this discrepancy is explained by the many phylogenies that are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay]. However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.

@@

The general workflow is described in FIGURE. Next, we will describe the technical details of each step.

@@

- The software MUSCLE [@edgar2004muscle] is used by default to perform sequence alignments. It has to be installed by the user, previous to running physcraper.
- First, all new sequences are aligned using default MUSCLE options.
- Then, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
- The final alignment is not further processed automatically. So, we encourage users
  to check it either by eye and perform manual refinement or using any of the many
  tools for alignment processing, to eliminate columns with no information.
- Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.
